---
title: "Chapter Four Homework"
author: "JJH"
date: "Last compiled on `r format(Sys.time(), '%d %B, %Y %H:%M')`"
output:
  html_document:
    toc: true
    toc_float: true
    keep_md: true
    self_contained: true
---

```{r, include = FALSE}
require(tidyverse)
require(car)
require(ALSM)
require(MASS)
```

| i:       | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8   | 9   | 10  |
|----------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|
| $X_{i}$: | 1   | 0   | 2   | 0   | 3   | 1   | 0   | 1   | 2   | 0   |
| $Y_{i}$: | 16  | 9   | 17  | 12  | 22  | 13  | 8   | 15  | 19  | 11  |

: Question 3.5 Airfreight breakage

```{r, include = TRUE}
airfreight <-
  tibble(
  x = c(1, 0, 2, 0, 3, 1, 0, 1, 2, 0),
  y = c(16, 9, 17, 12, 22, 13, 8, 15, 19, 11)
  )
```

```{r}
airfreight.lm <- lm(y ~ x, data = airfreight)
```

```{r}
airfreight.lm
```

```{r}
summary(airfreight.lm)
```

```{r}
Anova(airfreight.lm, type = "III")
```

**3.5** Refer to **Airfreight breakage** Problem 1.21.

d. Plot the residuals $e_{i}$ against $X_{i}$ to ascertain whether any departures from regression model (2.1) are evident. Wha tis your conclusion?

```{r}
plot(residuals(airfreight.lm) ~ airfreight$x,
     xlab = "Predictor", ylab = "Residual")
```

Plot of residuals against fitted values

```{r}
plot(residuals(airfreight.lm) ~ fitted.values(airfreight.lm),
     xlab = "Fitted value", ylab = "Residual")
```

**f. Prepare a time plot of the residuals. What information is provided by your plot?**

Sequence plot of residuals

```{r}
plot(airfreight.lm$residuals,
     type = "l",
     xlab = "Production run",
     ylab = "Residual")
abline(h = 0)
```

```{r}
par(mfrow = c(2, 2))
plot(airfreight.lm)
```

Plot of semistudentized residuals against fitted values

```{r}
plot(airfreight.lm$residuals/summary(airfreight.lm)$sigma ~ airfreight.lm$fitted.values, 
     xlab = "Fitted value", ylab = "Semistudentized residual")
```

Normal probability plot of residuals

```{r}
qqnorm(airfreight.lm$residuals,
       main = "Normal Q-Q plot of residuals") 
qqline(airfreight.lm$residuals)
```

**d.** Plot the residuals $\e_{i}$ against $X_{i}$ to ascertain whether any departures from regression model (21.) are evident. What is your conclusion?

```{r}
par(mfrow = c(2,2))
plot(airfreight.lm)
```

**f.** Prepare a time plot of the residuals. what information is provided by your plot?

**3.13** Refer to **Copier maintenance** Problem 1.20.

```{r}
copier <- read.csv("Copier.csv")
```

```{r}
copier
```

**a.** What are the alternative conclusions when testing for lack of fit of a linear regression function?


The alternative conclusions when testing for a lack of fit of a linear regression function are the following:

1. The regression of the function is not linear.
2. The error terms do not have a constant variance.
3. The error terms are not independent.
4. The model fits all but one or a few outlier observations.
5. The error terms are not normally distrusted.
6. One or several important predictor variables have been omitted from the model.

*Nonlinearity of Regression Function*
*Nonconstancy of error variance*
*Presence of Outliers*
*Nonindependence of error terms*

$H_{0}: E{Y} = \beta_{0} + \beta_{1} X$
$H_{a}: E{Y} \neq \beta_{0} + \beta_{1} X$

**b.** Perform the test indicated in part (a). Control the risk of Type I error at 0.05. State the decision rule and conclusion.

```{r}
copier.lm <- lm(Y ~ X, data = copier)
plot(Y ~ X, data = copier)
abline(copier.lm, col="red")
```

*F*-test for lack of fit

```{r}
anova(copier.lm, lm(Y ~ factor(X), data = copier))
```

```{r}
anova(copier.lm, data = copier)
```

**RSS** = *SSPE* = 2797.7
**Sum of Sq** = **SSLF** = 618.72

$F^{*} = (SSLF/c - 2) / (SSPE / n - c) = MSLF / MSPE$

$F^{*} = (618.719/8) / (2797.66 / 35) = 0.967557$

$F(0.95; 8; 35) = 2.21668$

If $F^{*} \leq 2.21668$, conclude $H_{0}$, otherwise $H_{a}$. Conclude $H_{0}$ 

$SSPE = \sum\sum (Y_{ij} - \hat Y_{j})^2$ 
$df = n - c$ 
$MSLF = SSLF/c - 2$ Number of groups minus 2

$SSLF = \sum\sum (\hat Y_{j} - \hat Y_{ij})^2$ 
$df = c - 2$ 
$MSPE = SSPE/n - c$ Total observations minus number of groups

**c.** **Sales growth.** A marketing researcher studied annual sales of a product that had been introduced 10 years ago. The data are as follows, where $X$ is the year (coded) and $Y$ is sales in thousands of units:

| i:        | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8   | 9   | 10  |
|-----------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|
| $X_{i}$:  | 0   | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8   | 9   |
| $Y_{i}$:  | 98  | 135 | 162 | 178 | 221 | 232 | 283 | 300 | 374 | 395 |

: Question 3.17 Sales growth

```{r, include = TRUE}
sales <-
  data.frame(
  x = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9),
  y = c(98, 135, 162, 178, 221, 232, 283, 300, 374, 395)
  )
```

```{r}
sales
```

**a.** Prepare a scatter plot of the data. Does a linear relation appear to be accurate?

```{r}
sales.lm <- lm(y ~ x, data = sales)
```

```{r}
plot(y ~ x, data = sales)
abline(sales.lm, col="red")
```

**b.** Use the Box-Cox procedure and standardization (3.36) to find an appropriate power transformation of $Y$. Evaluate $SSE$ for $\lambda = 0.3, 0.4, 0.5, 0.6, 0.7$.

What transformation of Y is suggested?

Transformations on Y

# Box-Cox transformation

```{r}
boxcox.sse(sales[,1],sales[,2],l=seq(-2,1,0.1))
```

| \lambda:|  0.3   | 0.4   | 0.5   | 0.6   | 0.7    | 
|---------|--------|-------|-------|-------|--------|
| $SSE$:  | 1099.7 | 967.9 | 916.4 | 942.4 | 1044.2 |


**c.** Use the transformation $Y' = \sqrt Y$ and obtain the estimated linear regression function for the transformed data.

```{r}
plot(y ~ x, data = sales)
```

```{r}
sales.lm_sqrt <- lm(sqrt(y) ~ x, data = sales)
```

```{r}
plot(sqrt(y) ~ x, data = sales)
```

```{r}
par(mfrow = c(2, 2))
plot(sales.lm_sqrt)
```

```{r}
summary(sales.lm_sqrt)
```
$Y = 1.07629X + 10.26093$

```{r}
plot(sales.lm$residuals ~ sales$x, xlab = "x", ylab = "Residual")
abline(h = 0)
```

```{r}
plot(sales.lm$residuals ~ sales$x, xlab = "x", ylab = "Residual")
```

```{r}
plot(sales.lm$residuals ~ sales.lm$fitted.values, 
     xlab = "Fitted value", ylab = "Residual")
```


**d.** Plot the estimated regression line and transformed data. Does the regression line appear to be a good fit to the transformed data?

```{r}
plot(sqrt(y) ~ x, data = sales)
```

**e.** Obtain the residuals and plot them against the fitted values. Also prepare a normal probability plot. What do your plots show?

Residuals

```{r}
sales.lm_sqrt$residuals 
```
Plot of residuals against fitted values

```{r}
plot(residuals(sales.lm_sqrt) ~ fitted.values(sales.lm_sqrt),
     xlab = "Fitted value", ylab = "Residual")
```

```{r}
fitted.values(sales.lm_sqrt)
```
```{r}
expected <- data.frame(actual = sales$y, predicted = predict(sales.lm_sqrt))
expected
```

```{r}
sales$Pred <- fitted(sales.lm_sqrt)
```

```{r}
sales$Pred
```
```{r}
sales$Resid <- sales$y - sales$Pred
```

```{r}
sales$Resid
```

```{r}
sales.sqrt_qq <- qqnorm(sales.lm_sqrt$residuals, main = "Normal Q-Q plot of residuals")
```


| $i$:              | 1     | 2     | 3     | 4     | 5     | 
|-------------------|-------|-------|-------|-------|-------|
| $e_{i}$:          | -0.36 | 0.28  | 0.31  | -0.15 | 0.30  |
| $\hat Y_{i}^{'}$: | 10.26 | 11.34 | 12.41 | 13.49 | 14.57 |
| *Expected value*: | -0.24 | 0.14  | 0.36  | -0.14 | 0.24  |

| $i$:              | 6     | 7     | 8     | 9     | 10    | 
|-------------------|-------|-------|-------|-------|-------|
| $e_{i}$:          | -0.41 | 0.10  | 0.47  | 0.47  | -0.07 |
| $\hat Y_{i}^{'}$: | 15.64 | 16.72 | 17.79 | 18.87 | 19.95 |
| *Expected value*: | -0.36 | 0.04  | -0.56 | 0.56  | -0.04 |

**f.** Express the estimated regression function in original units.

$Y = (1.07629X + 10.26093)^{2}$























