---
title: "Chapter Two HW"
author: "James J. Herlan"
editor: visual
---

```{r, include = FALSE}
require(tidyverse)
require(car)
require(ALSM)
```

| i:       | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8   | 9   | 10  |
|----------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|
| $X_{i}$: | 1   | 0   | 2   | 0   | 3   | 1   | 0   | 1   | 2   | 0   |
| $Y_{i}$: | 16  | 9   | 17  | 12  | 22  | 13  | 8   | 15  | 19  | 11  |

: Question 3.5 Airfreight breakage

```{r, include = TRUE}
airfreight <-
  tibble(
  x = c(1, 0, 2, 0, 3, 1, 0, 1, 2, 0),
  y = c(16, 9, 17, 12, 22, 13, 8, 15, 19, 11)
  )
```

```{r}
airfreight.lm <- lm(y ~ x, data = airfreight)
```

```{r}
airfreight.lm
```

```{r}
summary(airfreight.lm)
```

```{r}
Anova(airfreight.lm, type = "III")
```

**a. Estimate** $\beta_{1}$ with a 95 percent confidence interval. Interpret your interval estimate.

Confidence interval for slope

```{r}
confint(airfreight.lm, level = 0.95)
```

The estimated $\beta_{1}$ is 4.0, with a 95 percent confidence interval of 2.918388 $\leq$ $\beta_{1}$ $\leq$ 5.081612.

Based on the 95 percent interval, the researcher concludes that with 95% confidence, that the number of broken ampules increases somewhere between 2.92 and 5.08 for each time a carton is transferred.

**b. Conduct a** $t$ test to decide whether or not there is a linear association between number of times a carton is transferred ($X$) and number of broken ampules ($Y$). Use a level of significance of 0.05. State the alternatives, decision rule, and conclusion. What is the $P$-value of the test?

Critical value of $t$

```{r}
qt(0.975, 8)
```

p-value of t test

The p-value of t-test was calculated using the following:

```{r}
pt(8.528, 8, lower.tail = FALSE)*2
```

```{r}
(1 - pt(8.528, 8)) * 2
```

c.  $B_{0}$ represents here the mean number of ampules broken when no transfers of the shipment are made - i.e., when $X$ = 0. Obtain a 95 percent confidence interval for $B_{0}$ and interpret it.

d.  A consultant has suggested, on the basis of previous experience, that the mean number of broken ampules should not exceed 9.0 when transfers are made. Conduct an appropriate test, using $\alpha$ = 0.025. State the alternatives, decision rule, and conclusion.

Alternative hypotheses: $H_{A}$ is $\beta_{1} \neq 0$.

If $|t^{*}| \leq 2.306$ conclude $H_{0}$, otherwise $H_{A}$

The $p$-value is \< 0.05.

We reject the null hypothesis of $\beta_{1} = 0$.

c.  $B_{0}$ represents here the mean number of ampules broken when no transfers of the shipment are made - i.e., when $X$ = 0. Obtain a 95 percent confidence interval for $B_{0}$ and interpret it.

$b_{0}$ is 10.20. s{$b_{0}$} is 0.663.

```{r}
10.20 + (2.306*.663)
```

```{r}
10.20 - (2.306*.663)
```

**Answer:** 8.671122 $\leq$ $\beta_{0}$ $\leq$ 11.72888.

d.  A consultant has suggested, on the basis of previous experience, that the mean number of broken ampules should not exceed 9.0 when transfers are made. Conduct an appropriate test, using $\alpha$ = 0.025. State the alternatives, decision rule, and conclusion.

$H_{0}: \beta_{1} \leq 9$ $H_{A}: \beta_{1} > 9$

```{r}
(10.20 - 9)/0.663
```

$|t^{*}| = (10.20 - 9)/0.663 = 1.810$

If $|t^{*}| \leq 2.306$ conclude $H_{0}$, otherwise $H_{A}$

The p-value of t-test was calculated using the following:

```{r}
pt(1.810, 8, lower.tail = FALSE)
```

**2.15** Refer to the **Airfreight breakage** Problem 1.21.

| i:       | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8   | 9   | 10  |
|----------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|
| $X_{i}$: | 1   | 0   | 2   | 0   | 3   | 1   | 0   | 1   | 2   | 0   |
| $Y_{i}$: | 16  | 9   | 17  | 12  | 22  | 13  | 8   | 15  | 19  | 11  |

: Question 2.15 Airfreight breakage

**a.** Because of changes in airline routes, shipments may have to be transferred more frequently than in the past. Estimate the mean breakage for the following numbers of transfers: $X$ = 2, 4. Use separate 99 percent confidence intervals. Interpret your results.

CI for mean value of y at an x value

$X$ = 2

```{r}
predict(airfreight.lm, data.frame(x = 2), level = 0.99, interval = "confidence")
```

$X$ = 4

```{r}
predict(airfreight.lm, data.frame(x = 4), level = 0.99, interval = "confidence")
```

**b.** The next shipment will entail two transfers. Obtain a 99 percent prediction interval for the number of broken ampules for this shipment. Interpret your prediction interval.

```{r}
predict(airfreight.lm, data.frame(x = 2), level = 0.99, interval = "prediction")
```

**b.** In the next several days, three independent shipments will be made, each entailing two transfers. Obtain a 99 percent prediction interval for the mean number of ampules broken in the three shipments. Convert this interval into a 99 percent prediction interval for the total number of ampules broken in three shipments.

```{r}
ci.reg(airfreight.lm, data.frame(x = 2), type = "nm", alpha = 0.01, m = 3)
```

```{r}
3*(14.56543)
```

```{r}
3*(21.83457)
```

Answer: $44 \leq Total number of ampules broken \leq 65$

**d.** Determine the boundary values of the 99 percent confidence band for the regression line when $X_{h} = 2$ and when $X_{h} = 4$. Is your confidence band wider at these points than the corresponding confidence intervals in part (a)? Should it be?

```{r}
W2 <- 2*qf(0.99, 2, 8)
W2
```

```{r}
sqrt(W2)
```

Confidence band for entire regression line

```{r}
d <- data.frame(2)
ci.reg(lm(y~x, airfreight), d,  type = "s" , alpha = 0.01)
```

```{r}
d <- data.frame(4)
ci.reg(lm(y~x, airfreight), d,  type = "s" , alpha = 0.01)
```

**2.25** Refer to the **Airfreight breakage** Problem 1.21.

| i:       | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8   | 9   | 10  |
|----------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|
| $X_{i}$: | 1   | 0   | 2   | 0   | 3   | 1   | 0   | 1   | 2   | 0   |
| $Y_{i}$: | 16  | 9   | 17  | 12  | 22  | 13  | 8   | 15  | 19  | 11  |

: Question 2.25 Airfreight breakage

**a.** Set up the ANOVA table. Which elements are additive?

```{r}
Anova(airfreight.lm, type = "III")
```

**b.** Conduct an $F$ test to decide whether or not there is a linear association between the number of times a carton is transferred and the number of broken ampules; control the $\alpha$ risk at 0.05. State alternatives, decision rule, and conclusion.

Breakdown of SSTO

```{r}
boxplot(airfreight$y, airfreight.lm$residuals, col = "lightgray", main = "",
        names = c("y values", "residuals"))
```

```{r}
sd(airfreight$y)
```

```{r}
sd(airfreight.lm$residuals)
```

$H_{0}: \beta_{1} = 0, H_{a}: \beta_{1} \neq 0$.

$F^{*}$

```{r}
160.0/2.20
```

$F^{*} = 160.0/2.20 = 72.727$

$F(0.95; 1, 8) = 5.32$

```{r}
qf(0.95, 1, 8)
```

t critical value, squared

```{r}
(qt(0.975, 8))^2
```

If $F^{*} \leq 5.32$, then conclude $H_{0}$, otherwise $H_{a}$.

```{r}
1-pf(72.72727, 1, 8)
```

```{r}
(1-pt(8.528, 8))*2
```

**c.** Obtain a $t^{*}$ statistic for the test in part (b) and demonstrate numerically its equivalence to the $F^{*}$ statistic obtained in part (b).

```{r}
qf(0.95, 1, 8)
```

t critical value, squared

```{r}
(qt(0.975, 8))^2
```

$t^{*} = (4.00 - 0)/0.469 = 8.529$ $(t^{*})^{2} = (8.529)^{2} = 72.7 = F^{*}$

**d.** Calculate $R^{2}$ and $r$. What proportion of the variation in $Y$ is accounted for introducing $X$ into the regression model?

Coefficient of determination

```{r}
summary(airfreight.lm)$r.squared
```

Coefficient of correlation

```{r}
cor.airfreight = cor(airfreight$x, airfreight$y)
```

```{r}
cor.airfreight
```

```{r}
cor.airfreight^2
```
