---
title: "Chapter Four Homework"
author: "jjh"
editor: visual
---

```{r, include = FALSE}
require(tidyverse)
require(car)
require(ALSM)
require(MASS)
```

| i:       | 1   | 2   | 3   | 4   | 5   | 6   | 7   | 8   | 9   | 10  |
|----------|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|
| $X_{i}$: | 1   | 0   | 2   | 0   | 3   | 1   | 0   | 1   | 2   | 0   |
| $Y_{i}$: | 16  | 9   | 17  | 12  | 22  | 13  | 8   | 15  | 19  | 11  |

: Question 3.5 Airfreight breakage

```{r, include = TRUE}
airfreight <-
  tibble(
  x = c(1, 0, 2, 0, 3, 1, 0, 1, 2, 0),
  y = c(16, 9, 17, 12, 22, 13, 8, 15, 19, 11)
  )
```

```{r}
airfreight.lm <- lm(y ~ x, data = airfreight)
```

```{r}
airfreight.lm
```

```{r}
summary(airfreight.lm)
```

```{r}
Anova(airfreight.lm, type = "III")
```

**4.4** Refer to **Airfreight breakage** Problem 1.21.

Bonferroni 90% joint CIs for regression coefficients

```{r}
confint(airfreight.lm, level = 0.99)
```

**4.8** Refer to **Airfreight breakage** Problem 1.21.

**a.**

-   Simultaneous estimation of mean responses
-   Working-Hotelling procedure

```{r}
al = 0.05

(W = sqrt(2 * qf(1 - al, 2, nrow(airfreight) - 2)))

ci.reg(airfreight.lm, data.frame(x = c(0, 1, 2)), type = "w", alpha = al)
```

**b.**

-   Simultaneous estimation of mean responses
-   Bonferroni procedure

```{r}
g = 3

(B = qt(1-al/2/g, nrow(airfreight) - 2))

ci.reg(airfreight.lm, data.frame(x = c(0, 1, 2)), type = "b", alpha = al)
```

**c.**

-   Simultaneous prediction for new observations
-   Scheffe procedure

```{r}
al = 0.05

g = 3

(S = sqrt(g * qf(1 - al, g, nrow(airfreight) - 2)))

ci.reg(airfreight.lm, data.frame(x = c(0 ,1, 2)), type = "s", alpha = al)
```

**d.**

# Simultaneous prediction for new observations; Bonferroni procedure s

```{r}
ci.reg(airfreight.lm, data.frame(x = c(0, 1, 2)), type = "gn", alpha = al)
```
