---
title: "Chapter Six HW"
author: "James J. Herlans"
editor: visual
---

```{r, include = FALSE}
require(tidyverse)
require(car)
require(ALSM)
require(MASS)
require(nortest)
```

```{r, include = TRUE}
grocery <- read.csv('grocery.csv')
```

```{r}
grocery
```

**6.10**

**a** Fit regression model (6.5) to the data for three predictor variables. State the estimated regression function. How are $b_{1}$, $b_{1}$, $b_{1}$ here?

```{r}
grocery.lm <- lm(Y ~ X1 + X2 + X3, data = grocery)
```

```{r}
grocery.lm
```

```{r}
summary(grocery.lm)
```

**b.** Obtain the residuals and prepare a box plot of the residuals. What information does this plot provide?

```{r}
grocery.lm$residuals
```

```{r}
boxplot(grocery.lm$residuals)
```

```{r}
pairs(~grocery.lm$residuals+grocery.lm$fitted.values+X1+X2+I(X1*X2), data = grocery)
```

```{r}
plot(abs(grocery.lm$residuals)~grocery.lm$fitted.values, data = grocery)
```

**c.** Plot the residuals against $\hat Y$, $X_{1}$, $X_{2}$, $X_{3}$, and $X_{1}X_{2}$ on separate graphs. Also prepare a normal probability plot. Interpret the plots and summarize your findings.

```{r}
par(mfrow = c(2, 2))
plot(grocery.lm)
```

```{r}
qqnorm(grocery.lm$residuals) 
qqline(grocery.lm$residuals)
```


```{r}
shapiro.test(grocery.lm$residuals)
```

```{r}
lillie.test(grocery.lm$residuals)
```

**d.** Prepare a time plot of the residuals. Is there any indication that the error term are correlated?

```{r}
grocery.resid <- resid(grocery.lm)
```

```{r}
grocery.resid
```

```{r}
plot(grocery$X1, grocery.resid, 
     ylab = "Residuals", xlab = "X1", 
     main = "Grocery") 
abline(0, 0)                  # the horizon
```

```{r}
plot(grocery$X2, grocery.resid, 
     ylab = "Residuals", xlab = "X2", 
     main = "Grocery") 
abline(0, 0)                  # the horizon
```

```{r}
plot(grocery$X2, grocery.resid, 
     ylab = "Residuals", xlab = "X3", 
     main = "Grocery") 
abline(0, 0)                  # the horizon
```

*6.11* 

*a.* Test whether there is a regression relation, using level of significance 0.05. State the alternatives, decision rule, and conclusion. What does your result imply $B_{1}$, $B_{2}$, and $B_{3}$?
What is the $P$-value of the test?

```{r}
summary(grocery.lm)
```

```{r}
Anova(grocery.lm, type = 'III')
```

$H_{0}$: $B_{1}$ = $B_{2}$ = $B_{3}$

$H_{a}$: not all $B_{k}$ = 0 (*k* = 1, 2, 3).

$MSR$ = 725,535

$MSE$ = 20,531

$F^{*}$ = 725,535/531.9 = 35.337

$F(0.95; 3, 48)$ = 2.79806.

If $F^{*}$ < 2.79806 conclude $H_{0}$, otherwise $H_{a}$.

Conclude $H_{a}$. $P$-value = < 0.05.

**b.** Estimate $B_{1}$ and $B_{3}$ jointly by the Bonferroni procedure, using a 95 percent family confidence coefficient. Interpret your result.

```{r}
confint.lm(grocery.lm)
```

**c.** Calculate the coefficient of multiple determination $R^{2}$. How is this measure interpreted here?

```{r}
summary(grocery.lm)
```



